---
title: "D_P_Q_R of distributions"
author: "Dr K"
date: "September 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview of R's distribution functions

R has over 40 distribution functions 

### Description

Density, cumulative distribution function, quantile function and random variate generation for many standard probability distributions are available in the stats package.

### Details

The functions for the density/mass function, cumulative distribution function, quantile function and random variate generation are named in the form dxxx, pxxx, qxxx and rxxx respectively.  

### example using a normal distribution

#### Consider a normal distribution with mean 100 and standard deviaion 25  

we will create an array of x values from 0 to 200 with 2000 points in it.

```{r}
xval=seq(from=0,to=200,by=0.1)
```
Now we will apply dnorm to that array

```{r}
yvald=dnorm(xval,mean=100,sd=25)
```
Now let's plot the array 


```{r}
plot(xval,yvald)
```
  
  
### what does pnorm do

```{r}
yvalp=pnorm(xval,mean=100,sd=25)
```

```{r}
plot(xval,yvalp)
```

#### So p gives us the cumulative distribution function  

if we want to know the area under the curve to the left of x=103  

```{r}
pnorm(103,mean=100,sd=25)
```
### what does qnorm do ?  

qnorm gives us the quantiles we send in probabilities and it gives back x-values with those probabilities.   

suppose we want to know the x's where the probability is .25,.5,.75 and .8  

```{r}
probs=c(.25,.5,.75,.8)
qnorm(probs,mean=100,sd=25)
```
### finally what does rnorm do???

it takes a random sample of a specified size from the distribution  

```{r}
rsamp= rnorm(n=50,mean=100,sd=25)
hist(rsamp)
```

