---
title: "D_P_Q_R of distributions"
author: "Esteffani Maldonado"
date: "September 28, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview of R's distribution functions

R has over 40 distribution functions 

### Description

Density, cumulative distribution function, quantile function and random variate generation for many standard probability distributions are available in the stats package.

### Details

The functions for the density/mass function, cumulative distribution function, quantile function and random variate generation are named in the form dxxx, pxxx, qxxx and rxxx respectively.  

### example using a normal distribution

#### Consider a normal distribution with mean 100 and standard deviaion 25  

we will create an array of x values from 0 to 200 with 2000 points in it.

```{r}
xval=seq(from=0,to=200,by=0.1)
```
Now we will apply dnorm to that array

```{r}
yvald=dnorm(xval,mean=100,sd=25)
```
Now let's plot the array 


```{r}
plot(xval,yvald)
```
  
  
### what does pnorm do

```{r}
yvalp=pnorm(xval,mean=100,sd=25)
```

```{r}
plot(xval,yvalp)
```

#### So p gives us the cumulative distribution function  

if we want to know the area under the curve to the left of x=103  

```{r}
pnorm(103,mean=100,sd=25)
```
### what does qnorm do ?  

qnorm gives us the quantiles we send in probabilities and it gives back x-values with those probabilities.   

suppose we want to know the x's where the probability is .25,.5,.75 and .8  

```{r}
probs=c(.25,.5,.75,.8)
qnorm(probs,mean=100,sd=25)
```
### finally what does rnorm do???

it takes a random sample of a specified size from the distribution  

```{r}
rsamp= rnorm(n=50,mean=100,sd=25)
hist(rsamp)
```

## Now lets change from the normal distribution to the chi-squared distribution 
## there is just one parameter for such a distribution - df -- the degrees of freedom

```{r}
yvald=dchisq(xval,df=3) # 3 Deg of freedom
```
Now let's plot the array 


```{r}
plot(xval,yvald)
```

```{r}
yvald2=dchisq(xval,df=10) # 10 Deg of freedom
```

```{r}
plot(xval,yvald2)
```
```{r}
yvald3=dchisq(xval,df=30) # 30 Deg of freedom
```

```{r}
plot(xval,yvald3)
```

### we are going to plot Chi_Squared distributions degrees of Freedom go 3,10,15,25,30,50


```{r}
base=seq(from=0,to=100,by=.02)
chi3=dchisq(base,df=3)
chi10=dchisq(base,df=10)
chi15=dchisq(base,df=15)
chi25=dchisq(base,df=25)
chi30=dchisq(base,df=30)
chi50=dchisq(base,df=50)
plot(base,chi3,type="l")
title(main="Chi_Square plots \n Deg of freedom from 3 to 50")
lines(base,chi10,type="l",col=2)
lines(base,chi15,type="l",col=3)
lines(base,chi25,type="l",col=4)
lines(base,chi30,type="l",col=5)
lines(base,chi50,type="l",col=6)

```
  
## Now suppose the question Relates to a chisq distribution with 30 deg of freedom
## and the question is "at what value of x is the area under the curve (from the left up to the point) equal to .75?" 
## recall the area under any distribution function is always 1.

```{r}
plot(base,chi30,type="l",col=5)
abline(v=qchisq(p=.75,df=30))
print(qchisq(p=.75,df=30))
```

## Now suppose we want to know what area is to the left of the point 40 on the distribution function of chisq with 30 Deg of freedom.  

```{r}
plot(base,chi30,type="l",col=5)
abline(v=40)
print(pchisq(q=40,df=30))
```
## Now let us talk about the binomial distribution function, the whole idea goes like this -- we know the probability of getting say one bad apple, let's say that is .5 (or 50%).  Suppose we have a box of 100 apples, what's the probability of getting 0, 1, 2, ... ,99, 100 bad apples  


using and plotting binomial distributions

```{r}
par(mfrow=c(1,2))
x=0:100
p.5=dbinom(x,size=100,prob=.5)
plot(x,p.5,type="l",main="dbinom \n p=.5")
z.5=pbinom(x,size=100,prob=.5)
plot(x,z.5,type="l",main="pbinom \n p=.5")
par(mfrow=c(1,1))
```

Lets change p to .2

```{r}
par(mfrow=c(1,2))
x=0:100
p.2=dbinom(x,size=100,prob=.2)
plot(x,p.2,type="l",main="dbinom \n p=.2")
z.2=pbinom(x,size=100,prob=.2)
plot(x,z.2,type="l",main="pbinom \n p=.2")
par(mfrow=c(1,1))
```

Lets change p to .8

```{r}
par(mfrow=c(1,2))
x=0:100
p.8=dbinom(x,size=100,prob=.8)
plot(x,p.8,type="l",main="dbinom \n p=.8")
z.8=pbinom(x,size=100,prob=.8)
plot(x,z.8,type="l",main="pbinom \n p=.8")
par(mfrow=c(1,1))
```

let's compute the mean when p=.5

```{r}
x=0:100
p.5=dbinom(x,size=100,prob=.5)
data.5=x*p.5
mean.5=sum(data.5)
mean.5
```

let's compute the mean when p=.2

```{r}
x=0:100
p.2=dbinom(x,size=100,prob=.2)
data=x*p.2
mean.2=sum(data)
mean.2
```

how about the mean when p=.8

```{r}
x=0:100
p.8=dbinom(x,size=100,prob=.8)
data.8=x*p.8
mean.8=sum(data.8)
mean.8
```
## so a random variable with a size=n and prob=p  
## has a mean of n*p  

# how about variance or standard deviation?

```{r}
x=0:100
p.5=dbinom(x,size=100,prob=.5)
data.5=x*p.5
mean.5=sum(data.5)
#now center x by subtracting mean from each element
cx.5=x-mean.5
#and now sum the squared centered x's times their probabilities
var.5=sum(cx.5^2*p.5)
sd.5=sqrt(var.5)
var.5
sd.5
```
Let' see about our other two (p=.2 and p=.8)
```{r}
x=0:100
p.2=dbinom(x,size=100,prob=.2)
data.2=x*p.2
mean.2=sum(data.2)
#now center x by subtracting mean from each element
cx.2=x-mean.2
#and now sum the squared centered x's times their probabilities
var.2=sum(cx.2^2*p.2)
sd.2=sqrt(var.2)
var.2
sd.2
```

## now for p=.8

```{r}
x=0:100
p.8=dbinom(x,size=100,prob=.8)
data.8=x*p.8
mean.8=sum(data.8)
#now center x by subtracting mean from each element
cx.8=x-mean.8
#and now sum the squared centered x's times their probabilities
var.8=sum(cx.8^2*p.8)
sd.8=sqrt(var.8)
var.8
sd.8
```
## now check against these formula's

### $mean = n*p$

### $standard deviation = \sqrt(n*p*q)$




